%\section{Transcription}
\section{Other "omics" and "big data" integration}

\subsection{Transcriptor factor}
The promotor is where a transciptor factor binds.

Most genes of bacteria are regulated with the Lac repressor.

Genes are turned on and off by events (like it's too hot, you're poisoned by 
metal, etc).

A gene could have more of one promotor.

The are some information, not in the genom itself that is called epigeonimc 
that can do some modification to the organism (like caterpillar $\to$ 
butterfly).

SBML is a sort of formalism that allow you to describe gene, their position and 
other information. There are many computer programs that allow you to input 
these descriptions.

\section{Finding relevant variants from personal genomic data}

In 2008, 2 scientist did a lot of PCR reactions (36 million) in the genome of 
22 human tumor samples to get only the exome (45 million bases).

What do you do to process the genome?
Steps:
\begin{enumerate}
	\item Sequencing (Illumina proton solid) $\to$ FASTQ file
	\item Alignment (BWA many others) $\to$ SAM file
	\item Variant call (GATK\footnote{One variant per line} IonProton Caller) 
	$\to$ VCF file
	\item make sense (QueryOR some other)
\end{enumerate}

We want to know from any variant the percentage.

There are different annotation of all the variants. We can select the variants 
that are involved in the disease and that we know the causative genes.

\paragraph{QueryOR}
Flow:
\begin{enumerate}
	\item Upload VCF file
	\item Data processing from the QueryOR Server
	\item Interactive queries\footnote{See \url{http://queryor.cribi.unipd.it}}
\end{enumerate}

Sometime our gene is not know at all, and we cannot use QueryOR.