% Recap
If you have the whole human genome, you have 3 billion variants, with exon you 
have 14000 variants, that it's still a lot.

We want to know the probability of a disease for a new child from a family with 
a Father, a Mother and a Child with a genetic disease. Father and Mother are 
cousin.
\begin{enumerate}
	\item We take the exon and we search for variants. Even if one gene is not 
	working we have two genome, so there is the other that "save" the situation.
	If you have two cousin, you have to remember that they inherit 25\% of the 
	genome from the grandparent. If you inherit the variations from the same 
	ancestor you have a high probability that the genetic disease become 
	homozigote. So if you have consaguinion marriage it's very likely to have 
	genetic disease, because from two eterozigote you can have a homozigote 
	variant.
	\item ?? % he stopped explaining!!
\end{enumerate}

\subsection{GraphOne}
We have all these genes that are involved in a certain disease and we want find 
all the genes relates, from different sources (heterogeneous information). For 
this, you need a system that automatically learn (machine learning). This 
problem is solved with \texit{GraphOne}, that from a graph of association (that 
could be different association) and it transform the association in a table 
(called kernel). In this way, you transform the graph in something more simple 
(metrics). 
Based to the information that the system has, it will use different algorithm.

% Less track, but a more bit of logic (project).